---
title: "cross_validation"
author: "Rosie Kwon"
date: "2025-11-11"
output: html_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)
library(p8105.datasets)

```

```{r}
data("lidar")
```

Look at the data
```{r}
lidar

lidar_df =
  lidar |> 
  mutate(id = row_number())

lidar_df |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point()
```


## Create dataframes

```{r}
train_df = 
  sample_frac(lidar_df, size = .8) |> 
  arrange(id)

test_df = anti_join(lidar_df, train_df, by = "id")
```

Look at these
```{r}
ggplot(train_df, aes(x= range, y = logratio)) +
  geom_point() +
  geom_point(data = test_df, color = "red")
```

Fit a few models to `train_df`
```{r}
linear_mod = lm(logratio ~ range, data = train_df)
```

```{r}
train_df |> 
  add_predictions(linear_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```

`mgcv::gam` for non-linear models

```{r}
smooth_mod = mgcv::gam(logratio ~ s(range), data = train_df)
wiggly_mod = mgcv::gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df)
```

```{r}
train_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```
```{r}
train_df |> 
  add_predictions(wiggly_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```

Try computing our root mean squared errors (RMSEs) for each model.[prediction accuracy]
Smaller RMSE = better predictive accuracy.
```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

# Iterate!

`crossv_mc` preforms the training / testing split multiple times, a stores the datasets using list columns
```{r}
cv_df = 
  crossv_mc(lidar_df, n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Did this work? Yes
```{r}
cv_df |>  pull(train) |>  nth(3)
```

```{r}
cv_df |> 
  mutate(
    linear_fits = map(train, \(df) lm(logratio ~ range, data = df))
  )
```

Or I could make a function
```{r}
lidar_lm = function(df){
  lm(logratio ~ range, data = df)
}

cv_df = 
  cv_df |> 
  mutate(
    linear_fits = map(train, lidar_lm),
    smooth_fits = map(train, \(df) mgcv::gam(logratio ~ s(range), data = df)),
    wiggly_fits = map(train, \(df) mgcv::gam(logratio ~ s(range, k = 50), sp = 10e-8, data = df))
  ) |> 
  mutate(
    rmse_linear = map2_dbl(linear_fits, test, rmse),
    rmse_smooth = map2_dbl(smooth_fits, test, rmse),
    rmse_wiggly = map2_dbl(wiggly_fits, test, rmse)
  )
```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x= model, y = rmse)) +
  geom_violin()
```

